---
title: "Metabolite Dendrogram"
author: "Jeremy Ash, Melanie Odenkirk"
output: 
  html_document:

    theme: united  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = T)
library(httr)
library(jsonlite)
library(fingerprint)
library(rcdk)
library(ggplot2)
library(ggtree)
library(ape)
library(phangorn)
library(limma)
library(digest)
library(pheatmap)
library(dplyr)
library(RColorBrewer)

knitr::opts_knit$set(root.dir = 'data/')
```


# Healthy versus Pregnancy Disorders, all lipids

## Statistical Analysis

Lipids were assessed for statistical significance in a Disease (GDM or PRE) versus control comparison. Here, an alpha cut off of 0.05 was used with a Dunnett multiple testing correction. More information on this process is given in the methods section. 


## Building p-value heatmaps of statistically significant lipids and parsing lipids by fatty acyl annotation
```{r, message=FALSE, warning=FALSE}
# Combine together data for healty verus control comparison
# Combine negative and positve data sets

pd_fn <- c("PRE vs. Ctrl", "GDM vs. Ctrl")

fn <- pd_fn[1]
sig.df <- data.frame(matrix(nrow = 319, ncol = length(pd_fn)))
for (i in 1:length(pd_fn)) {
  fn <- pd_fn[i]
  df1 <- read.csv(paste0(fn, "__Negative.csv"))
  df2 <- read.csv(paste0(fn, "__Positive.csv"))
  df <- rbind(df1, df2)
  sig.df[, i] <- (df$P_value_T_PD_vs_Ctrl* df$Flag_PD_vs_Ctrl) 
  colnames(sig.df)[i] <- fn
}

sig.df$Name <- df$Name
sig.df <- sig.df[c(3, 1:2)]

lipids <- sig.df$Name
sig.df$Name <- gsub("/", "_", sig.df$Name)

#Drop the Bs and keep the As for now
sig.df$Name <- gsub("_[A]$", "", sig.df$Name)
sig.df <- sig.df[!grepl("_[B]$", sig.df$Name), ]

# Split the rows with multiple potential lipids ids into 
# separate entries
for (i in 1:nrow(sig.df)) {
  lipids <- unlist(strsplit(sig.df$Name[i], ";"))
  if (length(lipids) > 1) {
    r <- sig.df[i, ]
    sig.df <- sig.df[-i, ]
    mat <- data.frame(matrix(NA, nrow = length(lipids), ncol = ncol(sig.df)))
    colnames(mat) <- colnames(sig.df)
    for (j in 1:length(lipids)) {
      mat[j, ] <- data.frame(Name = lipids[j], r[-1])
    }
    mat$Name <- lipids
    sig.df <- rbind(sig.df, mat)
  }
}
#Need to do this twice
for (i in 1:nrow(sig.df)) {
  lipids <- unlist(strsplit(sig.df$Name[i], ";"))
  if (length(lipids) > 1) {
    r <- sig.df[i, ]
    sig.df <- sig.df[-i, ]
    mat <- data.frame(matrix(NA, nrow = length(lipids), ncol = ncol(sig.df)))
    colnames(mat) <- colnames(sig.df)
    for (j in 1:length(lipids)) {
      mat[j, ] <- data.frame(Name = lipids[j], r[-1])
    }
    mat$Name <- lipids
    sig.df <- rbind(sig.df, mat)
  }
}

for (i in 1:nrow(sig.df)) {
  lipids <- unlist(strsplit(sig.df$Name[i], ";"))
  if (length(lipids) > 1) {
    r <- sig.df[i, ]
    sig.df <- sig.df[-i, ]
    mat <- data.frame(matrix(NA, nrow = length(lipids), ncol = ncol(sig.df)))
    colnames(mat) <- colnames(sig.df)
    for (j in 1:length(lipids)) {
      mat[j, ] <- data.frame(Name = lipids[j], r[-1])
    }
    mat$Name <- lipids
    sig.df <- rbind(sig.df, mat)
  }
}

for (i in 1:nrow(sig.df)) {
  lipids <- unlist(strsplit(sig.df$Name[i], ";"))
  if (length(lipids) > 1) {
    r <- sig.df[i, ]
    sig.df <- sig.df[-i, ]
    mat <- data.frame(matrix(NA, nrow = length(lipids), ncol = ncol(sig.df)))
    colnames(mat) <- colnames(sig.df)
    for (j in 1:length(lipids)) {
      mat[j, ] <- data.frame(Name = lipids[j], r[-1])
    }
    mat$Name <- lipids
    sig.df <- rbind(sig.df, mat)
  }
}

sig.df <- unique(sig.df)

duplicated(sig.df$Name)

sig.df <- sig.df[-c(180, 181, 182, 184, 185, 186, 187, 188, 189, 190, 191, 193, 194, 235), ]

rownames(sig.df) <- sig.df$Name

sig.df[sig.df == 0] <- NA

ab.set <- sig.df$Name
ab.set <- sub(".*\\(" , "", ab.set)
ab.set <- sub("\\).*" , "", ab.set)
ab.set <- sub(".*-" , "", ab.set)
ab.set <- sub("d" , "", ab.set)
ab.set <-strsplit(ab.set, split = "_")
ab.set <- unlist(ab.set)
ab.set <- unique(ab.set)
names(ab.set) <- c("group")
FA <- c("0:0", "12:0", "14:0", "15:0", "16:0", "16:1", "17:0", "18:0", "18:1", "18:2", "18:3", "20:0", "20:1", "20:3", "20:4", "22:0", "22:4", "22:6")
ab.set <- ab.set[order(match(ab.set, FA))]

lipid.set <- data.frame(stringsAsFactors = F)
l <- ab.set[1]
for(l in ab.set) {
  tmp <- sig.df$Name[grepl(l, sig.df$Name)]
  lipid.set <- rbind(lipid.set, cbind(tmp, rep(l, length(tmp))))
}

colnames(lipid.set) <- c("lipids", "group")
length(unique(lipid.set$group))
table(lipid.set$group)

highfreq.groups <- unique(lipid.set$group)[table(lipid.set$group) > 10]

lipid.set.trim <- lipid.set[lipid.set$group %in% highfreq.groups, ]
table(droplevels(lipid.set.trim$group))
lipid.set.trim<- filter(lipid.set.trim, group != "8:0")

sig.df.full <- merge(lipid.set.trim, sig.df, by.x = "lipids", by.y = "Name")

head(sig.df.full)

sig.df.full <- sig.df.full[order(sig.df.full$group), ]
sig.df.mat <- sig.df.full[, -(1:2)]
sig.df.mat <- data.matrix(sig.df.mat)

group <- data.frame(sig.df.full$group)
group <- droplevels(group)
table(group)

length(group)

row.names(group) <- paste0(sig.df.full$lipids, sig.df.full$group, sep = ".")
row.names(sig.df.mat) <- paste0(sig.df.full$lipids, sig.df.full$group, sep = ".")

dim(sig.df.mat)
length(group)
```


```{r}
breaks <- seq(from=min(range(-0.05)), to=max(range(0.05)), length.out=1000)
midpoint <- which.min(abs(breaks - 0))
rampCol1 <- colorRampPalette(c("#00e8f0","#0d78ca", "#002060"))(midpoint)
rampCol2 <- colorRampPalette(c( "#600000", "red", "#FF4D4D"))(1000-(midpoint))
rampCols <- c(rampCol1,rampCol2)

pheatmap(sig.df.mat,color = rampCols, breaks = breaks, annotation_row = group, gaps_col ="1", cellwidth = 50, cellheight = 10, treeheight_row = 100, treeheight_col = 100,  cluster_row = F, cluster_col = F, fontsize = 10, legend = T, filename = "FA clustering_tail length order_ALL_Lipids.tiff") 

sig.df.nonzero <- sig.df.mat[!(rowSums(is.na(sig.df.mat))==NCOL(sig.df.mat)),] 

pheatmap(sig.df.nonzero, color = rampCols, breaks = breaks,annotation_row = group, gaps_col ="1", cellwidth = 50, cellheight = 10, treeheight_row = 100, treeheight_col = 100,  cluster_row = F, cluster_col = F, fontsize = 10, legend = T, filename = "FA clustering_tail length order_SIG_Lipids.png")

```


